# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








Standardizing a variable means subtracting the mean, and then dividing by the standard deviation. Let’s use a loop to standardize the numeric columns in the [Western Collaborative Group Study](https://clinicaltrials.gov/ct2/show/NCT00005174). This study began in 1960 with 3154 men ages 39-59, who were employed in one of 11 California based companies. They were followed until 1969 during this time, 257 of these men developed coronary heart disease (CHD). You can read this data in with the code below. You can access this dataset with the following code:

```{R}

suppressMessages(library(foreign))
wcgs <- read.dta("https://drive.google.com/uc?export=download&id=0B8CsRLdwqzbzYWxfN3ExQllBQkU")
```

The data has the following variables:



WCGS has the following variables:

-----------------------------------------------------------
Name    Description
------- -------------------------------------------
id      Subject identification number

age     Age in years

height  Height in inches

weight  Weight in lbs.

sbp     Systolic blood pressure in mm 

dbp     Diastolic blood pressure in mm Hg

chol    Fasting serum cholesterol in mm 

behpat  Behavior

  1       A1

  2       A2

  3       B3

  4       B4

ncigs   Cigarettes per day

dibpat  Behavior

1       type A

2       type B

chd69   Coronary heart disease

1       Yes

0       no

typechd Type of CHD

1       myocardial infarction or death

2       silent myocardial infarction

3       angina perctoris

time169 Time of CHD event or end of follow-up

arcus   Arcus senilis

0       absent

1       present

bmi     Body Mass Index
-----------------------------------------------------------




### Question 1: Standardize Function

A. Create a function called standardize.me() that takes a numeric vector as an argument, and returns the standardized version of the vector. 

```{r }
x <- c(1:10)
standardize.me <- function(x){ #created a function called standardize.me() with an argument that assigns x as numeric
  x <-  (x - mean(x, na.rm = TRUE)) /sd(x, na.rm = TRUE) #set the body of the function to return a standardized version of the vector where I subtract the mean of x and divided by the standard deviation of x
  return(x) #this outputs x as a standardized function
}

#args(standardize)
```
B. Assign all the numeric columns of the original WCGS dataset to a new dataset called WCGS.new.
```{r }
library(dplyr) #called library dplyr
WCGS.new <- select(wcgs, -id, -behpat, -dibpat, -chd69, -typchd69, -arcus, -agecat) #used the dplyr function "select" to select all numeric columns in the original WCGS dataset by removing the indicated columns and assigned them to the new dataset "WCGS.new"
```

C. Using a loop and your new function, standardize all the variables WCGS.new dataset.
```{r }
for (i in colnames (WCGS.new)){ #assigned the variable i to all the variables in the column of the WCGS.new dataset
  WCGS.new[, i] <- standardize.me(WCGS.new[,i] ) #I called the newly generated standardized.me function on the colnames of i and performed an over-write of standarized dataset. 
  }

```
D. What should the mean and standard deviation of all your new standardized variables be? Test your prediction by running a loop
```{r }
for (i in colnames (WCGS.new)) {#Used the for loop to assign i to variables in the colnames of the WCGS.new dataset
  WCGS.new_mean_sd <- c(mean(WCGS.new [,i], na.rm = TRUE),  sd(WCGS.new [,i], na.rm = TRUE)) #used the cantenation sign to combine the commands for finding the mean and standard deviation for iteration of all the numeric colnames in the WCGS.new dataset. Wraped the commands to a vector name.
  print(WCGS.new_mean_sd) #printed the new vector to see all the means and sds for all the columns of the iterated dataset
}

```




### Question 2: Looping to Calculate

A. Using a loop, calculate the mean weight of the subjects separated by the type of CHD they have.
```{r }
mean_weight <- for (i in wcgs$typchd69) { #used the for loop to assign i to all variables of the typchd69 column in the "wcgs" dataset. Assigned this to a vector called "mean_weight"
  mean(wcgs[,"weight"], na.rm = TRUE) #attempted to find the mean of the weight column
  print(mean_weight) #printed out the vector name "mean_weight"
  stop() #since there was an error in the code that caused it to continue running, I used the "stop()" function to terminate the run
  
}

```
B. Now do the same thing, but now don’t use a loop
```{r }
library(dplyr) #called the dplyr library
wcgs %>% #called the wcgs dataset 
  select(id, weight, typchd69)%>% #used the dplyr select function to select the variables "id", "weight" and "typch69"
  group_by(typchd69)%>% #grouped_by typchd69
  summarize(mean_weight = mean(weight))%>% #summarized the mean of the weight for the subjects according to their type of cdh
  collect() #used the dplyr collect function to retrieve the data into a local tibble
  


```

